\documentclass[main.tex]{subfiles}


\begin{document}


\section{Acknowledgements}

This research was supported by National Institute of Allergy and Infectious Diseases grant DP2 AI175475 (S.R.); National Science Foundation Graduate Research Fellowship Program grants 1745302 (M.F.A.), 2140743 (J.A.), and 2141064 (M.F.A.); and National Institutes of Health grant T32GM145407 (J.A.).
We thank Miriam L. Rittenberg and Mateo Valenzuela for assistance with experiments.


\section{Author Contributions}

S.R. and M.F.A. conceived the project.
M.F.A. performed the experiments with SARS-CoV-2 segments.
J.A. performed the experiments with other coronavirus segments.
J.P. performed the experiments with TGEV-infected ST cells.
M.F.A. wrote SEISMIC-RNA with contributions from S.G., Y.J.M.T., A.L., and J.A.
M.F.A. analyzed the data with contributions from J.A.
M.F.A. drafted the manuscript.
All authors reviewed the manuscript and provided comments.


\section{Ethics Declarations}

The authors declare no competing interests.


\section{Data Availability}

All sequencing data generated in this study have been deposited into the NCBI Short Read Archive under accession code \href{https://www.ncbi.nlm.nih.gov/bioproject/PRJNA1103196}{PRJNA1103196}.


\section{Code Availability}

Documentation for SEISMIC-RNA, including instructions for installation, is hosted on GitHub Pages: \url{https://rouskinlab.github.io/seismic-rna}.
Source code for SEISMIC-RNA is available from GitHub: \url{https://github.com/rouskinlab/seismic-rna}. 
Shell scripts for running SEISMIC-RNA, auxiliary scripts for data analysis, supplementary files, and LaTeX source code for this manuscript are also available from GitHub: \url{https://github.com/rouskinlab/search-map}.


\end{document}
